---
title: "Reproducible Result Assignemnt1"
author: "Abhinav"
date: "19 May 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Set up the environment
```{r setup_environment, include=FALSE}
library(dplyr)
library(lattice)
getwd()
```
#1. Read data

``` {r READ_data, echo=TRUE}
 DF<- read.csv("activity.csv")
 str(DF)
 head(DF)
```

##Analyze the data
``` {r Analyze_data, echo=TRUE}
 Steps_Data <- summarise(group_by(DF, date), Total_Daily_Steps=sum(steps) ,n())
 head(Steps_Data)
 
```

#2. Histogram- Number of steps per day
``` {r HIstogram_Number_Steps_Per_DAY, echo=TRUE} 
 x <-as.numeric(na.omit(Steps_Data$Total_Daily_Steps))
 hist(x, breaks=10, main= "Total Number of Steps Taken per Day",col="green", xlab= "Total No of Steps",axes= TRUE)
```

#3. Mean, Median No of Steps
``` {r MEAN_MEDIAN, echo=TRUE}
mean(Steps_Data$Total_Daily_Steps,na.rm= TRUE)
median(Steps_Data$Total_Daily_Steps,na.rm= TRUE)
``` 

#4. Average Daily Activity Pattern- Average number of steps taken
``` {r Average_steps, echo=TRUE}
avg_step <- aggregate(steps ~ interval, data = DF, mean, na.rm = TRUE)
plot(avg_step$interval, avg_step$steps, type = "l", lwd = 2, col = "navy",
     main = "Time Series: Average Number of Steps Taken", axes = TRUE,
     xlab = "5-minute interval", ylab = "Average number of steps")
```

#5.  Which 5 Min interval contains maximum number of steps?
``` {r Max_steps_per_5min_Interval, echo=TRUE}
avg_step$interval[which.max(avg_step$steps)]
```

#Imputing missing values
## Total number of missing values
``` {r Number_NA, echo=TRUE}
sum(is.na(DF))
```

#6. Strategy for filling in null values - 
##Replace missing data with mean of that column, use mean of 5 min interval corresponding to missing value. Create a new dataset with missing values filled in
``` {r Filling_in_null_values, echo=TRUE}
New_DF <- DF

for(i in 1:nrow(DF))    
  {
    if (is.na(New_DF[i,1]))
     {
       int_5= New_DF[i,3]
       New_DF[i,1] <- mean(New_DF$steps[New_DF$interval== int_5],na.rm=TRUE)
     }
  }
 head(New_DF)
```

#7. Histogram of the total number of steps taken each day and report the mean and median total number of steps taken per day.
``` {r Total_number_of_steps_taken_each_day, echo=TRUE}
New_x <- summarise (group_by(New_DF, date), Total_Daily_Steps=sum(steps) ,n())
  
  a <- as.numeric(New_x$Total_Daily_Steps)
  hist(a,breaks=10, main='Total number of steps taken each day-NA replaced',   xlab='Steps', ylim=c(0,30),axes= TRUE, col= "green")
```

##Mean, Median No of Steps
``` {r Mean_Median_Steps, echo=TRUE}
mean(New_x$Total_Daily_Steps,na.rm= TRUE)
median(New_x$Total_Daily_Steps,na.rm= TRUE)
```


##New factor variable in the dataset with two levels - "weekdays" and "weekend" indicating whether a given date is a weekday or weekend day.
``` {r New_factor_variable, echo=TRUE}
New_DF$Day_Status <- ifelse(weekdays(as.Date(New_DF$date)) %in% c("Saturday", "Sunday"), "weekend", "weekday")
```

#8.Time series plot (i.e. type = "l") of the 5- minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
``` {r Time_series_plot, echo=TRUE}
avg_step <- aggregate(steps ~ interval + Day_Status, data = New_DF, mean, na.rm = TRUE)

xyplot(steps ~ interval| Day_Status, avg_step, 
       type="l", 
       lwd=2, 
       xlab="Interval", 
       ylab="Number of steps", 
       layout=c(1,2))
```

